---
title: "symbiosis growth analysis"
author: "Analysis - David J. Weston<br/>Experimentation - Colleen Husley & Megan Gable"
date: "11/16/2018"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}

library(captioner)
fig_nums <- captioner(prefix = "Fig.")
fig.1_cap <- fig_nums(name = "fig_1", 
                        caption = "2018 data, change in growth (initial weight (wt) - final wt among the two symbiotic partners alone and when added togather but grown seperately (_seperate) and when grown together in same well (_symbiosis). Cyano in symbiosis includes, endophytes, epiphytes and free living cyano.  Data incuded all pH conditions")
fig.2_cap <- fig_nums(name = "fig_2", 
                        caption = "2018 data, change in growth among interacting partners faceted by starting pH condiiton. Measured components as described in fig. 1")
fig.3_cap <- fig_nums(name = "fig_3", 
                        caption = "2018 data, total well biomass change from experiment start to finish. Includes moss + cyano endophyte and free living cyanobacteria not attached to plant")

fig.4_cap <- fig_nums(name = "fig_4", 
                        caption = "2018 data, change in pH change as measured by (final - initial values)")

fig.5_cap <- fig_nums(name = "fig_5", 
                        caption = "2018 data, absolute value of pH change as measured by (final - initial values)")
fig.6_cap <- fig_nums(name = "fig_6", 
                        caption = "2017 data, change in growth (initial weight (wt) - final wt among the two symbiotic partners alone, when added togather but grown seperately (_seperate) and when grown together in same well (_symbiosis). Conditions aggregated over pH") 
fig.7_cap <- fig_nums(name = "fig_7", 
                        caption = "2017 data, change in growth (initial weight (wt) - final wt among the two symbiotic partners alone, when added togather but grown seperately (_seperate) and when grown together in same well (_symbiosis). Similar to Fig. 6 except faceted by pH")
fig.8_cap <- fig_nums(name = "fig_8", 
                        caption = "Combined 2018 and 2017 data for change in growth (initial weight (wt) - among the two symbiotic partners alone, when added togather but grown seperately (_seperate) and when grown together in same well (_symbiosis). Data is faceted by pH")


```

# Introduction
The symbiosis between Sphagnum mosses and cyanobacteria has long been recognized (Limpricht, 1890) and has been implicated in improved moss growth. This together with evidence form other plant - cyanobacteria symbioses suggest that this is not just a symbiosis (interacting species), but perhaps a mutualsitc interaction -- thereby benefiting both interacting speices. To investigate this further, Colleen Hulsey and Megan Gable conducted a numer of experiments to investigate the growth consequences of cyanobacteria and sphagnum symbiosis on growth patterns at differing pH. conditions.   

# 2018 Experimental approach and measurements

*Experimental conditions & design
  +conducted on 08/09/2018
  +plants pH nitrogen depleted for 2 weeks prior to start
  +experiment ran for 2 weeks
  +condition 1: Nitrogen (N), 6 replicates -N, 2 replicates for low N (5mL/L NaNO3-- Mm?)
  +Condition 2: pH, 3.5, 5.5, 8.5
  +speices: Sphagnum fallax MN alone, Nostoc muscorum 1037 alone, both species together
  +amounts: 1 moss capitulum, 0.1g cyanobacteria
  +3 - 12 well plates, 2 ml liquid medium per well. 
  
*Measurements
  +fluorcam: area_mm2, QY_max (initial, weekly, final)
  +pH initial, pH weekly, pH final
  +weight beginning and end after water removed with centrifugation

*tests
  +simple scatter and boxplots to view realtionship


# 2018 Data import and summary

```{r, warning = FALSE, message = FALSE}
library(tidyverse)
library(dplyr)
#knitr by default will set the working directory to the source file location
#setwd("~/Documents/Manuscripts:data/Symbiosis_study_2018")
dat1 <- read_csv("fallax_cyano_symbiosis_colleen_10_01_2018.csv") %>% mutate(delta_weight = delta_weight_g *1000)
names(dat1)
summary(dat1)
dim(dat1)

```

# Boxplot analysis for 2018 data
  1. Delta growth (weight) by organism regardless of pH; 2. Delta growth by organism and by pH; 3. Delta growth for entier wel by pH
  
```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.1_cap}

ggplot(dat1, aes(color = component_measured,y= delta_weight, x= component_measured)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title=element_text(size=16,face="bold"),
        legend.text=element_text(size=16))

```


```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.2_cap}

t<-ggplot(dat1, aes(color = component_measured, y= delta_weight, x= component_measured)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title=element_text(size=16,face="bold"),
        legend.text=element_text(size=16))
t+facet_grid(.~initial_pH)

```


```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.3_cap}

dat.well<-dat1 %>% filter(component_measured == "Cyanobacteria_moss_symbiosis")
t<-ggplot(dat.well, aes(y= delta_weight, x= component_measured)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        legend.text=element_text(size=16))
t+facet_grid(.~initial_pH)

```


```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.4_cap}

dat.ph.change<-dat1 %>% mutate(pH_change = final_pH - initial_pH)

t<-ggplot(dat.ph.change, aes(color = component_measured, y= pH_change, x= component_measured)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.text=element_text(size=16))
t+facet_grid(.~initial_pH)



```



```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.5_cap}
dat.ph.abs.change<-dat1 %>% mutate(abs_pH_change = abs(final_pH - initial_pH))

t<-ggplot(dat.ph.abs.change, aes(color = component_measured, y= abs_pH_change, x= component_measured)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        legend.text=element_text(size=16))
t+facet_grid(.~initial_pH)

```


# 2017 Experimental approach and measurements

need experimental details here...





# 2017 Data import and summary

```{r, warning = FALSE, message = FALSE}

dat2 <- read_csv("fallax_cyano_symbiosis_allyssa_colleen_2017.csv")
dat3 <- dat2 %>% filter(pH == 3.5 | pH == 5.5 | pH == 8.5)
names(dat3)
summary(dat3)
dim(dat3)

```


# Boxplot analysis of 2017 data

```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.6_cap}

ggplot(dat3, aes(color = condition,y= delta_weight, x= condition)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=16,face="bold"),
        legend.text=element_text(size=16))


```


```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.7_cap}

t<-ggplot(dat3, aes(color = condition,y= delta_weight, x= condition)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title=element_text(size=16,face="bold"),
        legend.text=element_text(size=16))
t+facet_grid(.~pH)

```




```{r, echo=FALSE, fig.width = 15, fig.height=10, fig.align='center',fig.cap = fig.8_cap}

dat.2018 <- dat1 %>% filter(component_measured == "moss_alone" | component_measured == "cyano_alone" | component_measured == "Cyanobacteria_moss_symbiosis" |
                              component_measured =="Cyanobacteria_moss_seperate") %>%
  select(condition = component_measured, pH = initial_pH, delta_weight) %>%
  mutate(condition = paste0(condition, "_2018"))

dat.2017 <- dat3 %>% select(condition, pH, delta_weight) %>%
  mutate(condition = paste0(condition, "_2017"))

dat.combined<-rbind(dat.2018, dat.2017)


t<-ggplot(dat.combined, aes(color = condition,y= delta_weight, x= condition)) +
  geom_boxplot(position=position_dodge(.8)) +
  geom_jitter(position=position_dodge(.8)) +
  theme(axis.text=element_text(size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        axis.title=element_text(size=16,face="bold"),
        legend.text=element_text(size=16))
t+facet_grid(.~pH)



```


# ANOVA analysis of 2017 and 2018 data.

Specifically testing if cyano and moss grown alone, increased in growth at a different rate than when grown together in smbiosis.


```{r, echo=FALSE}

# converting pH and condition to factor

dat.2017$pH<-factor(dat.2017$pH, levels =c(3.5, 5.5, 8.5), labels = c("pH_3.5", "pH_5.5", "pH_8.5"))
dat.2017$condition<-factor(dat.2017$condition, levels =c("cyano_alone_2017", "moss_alone_2017", "Cyanobacteria_moss_seperate_2017", "Cyanobacteria_moss_symbiosis_2017"), labels = c("cyano_alone_2017", "moss_alone_2017", "Cyanobacteria_moss_seperate_2017", "Cyanobacteria_moss_symbiosis_2017"))

#check with str(dat.2017)

dat.2017.aov2<-aov(delta_weight ~ condition + pH + condition:pH, data = dat.2017)
summary(dat.2017.aov2)

```

computing summary stats

```{r, echo=FALSE}

dat.2017 %>% group_by(condition, pH) %>% summarize(count = n(), mean = mean(delta_weight, na.rm = TRUE), sd = sd(delta_weight, na.rm = TRUE))

```


mean seperation iwth Tukeys

```{r, echo=FALSE}
TukeyHSD(dat.2017.aov2, which = "condition")

```

resdiual check

```{r, echo=FALSE}

par(mfrow = c(1,2))
plot(dat.2017.aov2, 1)
plot(dat.2017.aov2, 2)

```

```{r, echo=FALSE}
# Extract the residuals
aov_residuals <- residuals(object = dat.2017.aov2)
# Run Shapiro-Wilk test
shapiro.test(x = aov_residuals)

```


Starting the 2018 data

```{r, echo=FALSE}

# converting pH and condition to factor

dat.2018$pH<-factor(dat.2018$pH, levels =c(3.5, 5.5, 8.5), labels = c("pH_3.5", "pH_5.5", "pH_8.5"))
dat.2018$condition<-factor(dat.2018$condition, levels =c("moss_alone_2018", "cyano_alone_2018", "Cyanobacteria_moss_symbiosis_2018", "Cyanobacteria_moss_seperate_2018"), labels = c("moss_alone_2018", "cyano_alone_2018", "Cyanobacteria_moss_symbiosis_2018", "Cyanobacteria_moss_seperate_2018"))



#check with str(dat.2018)

dat.2018.aov2<-aov(delta_weight ~ condition + pH + condition:pH, data = dat.2018)
summary(dat.2018.aov2)

```

computing summary stats for 2018

```{r, echo=FALSE}

dat.2018 %>% group_by(condition, pH) %>% summarize(count = n(), mean = mean(delta_weight, na.rm = TRUE), sd = sd(delta_weight, na.rm = TRUE))

```


mean seperation with Tukeys for 2018

```{r, echo=FALSE}

TukeyHSD(dat.2018.aov2, which = "condition")

```

resdiual check 2018 data

```{r, echo=FALSE}
par(mfrow = c(1,2))
plot(dat.2018.aov2, 1)
plot(dat.2018.aov2, 2)

```

Shapiro-Wilk test for normality

```{r, echo=FALSE}
# Extract the residuals
aov_residuals <- residuals(object = dat.2017.aov2)
# Run Shapiro-Wilk test for normatily of residuals
shapiro.test(x = aov_residuals)


```


```{r}
#Levene's Test for homogenity of variances

library(car)
leveneTest(delta_weight ~ condition, data = dat.2018)
leveneTest(delta_weight ~ pH, data = dat.2018)

```


